\subsubsection{The Zone Language}
\label{sec:zonelang}

Our statically typed language is inspired by array and data flow
programming languages such as Fortran\cite{Fortran}, APL\cite{APL}, and
LINQ\cite{LINQ} where one expresses computation based on operations on
vectors and arrays. In our language, one can define two vectors of size
\texttt{100} by

\begin{verbatim}
n :: Integer = 1000;
as :: []Real = rand.Real(n);
bs :: []Real = rand.Real(n); 
\end{verbatim}


The syntactic style for ZOne is borrowed from a other languages
	which we think have good syntax representations.
The type declaration position was browed from Go\cite{golang},
	the function declarations from Dart\cite{dartlang},
	the type notation from Julia\cite{julialang},
	and the semantics from array languages such as Fortran\cite{Fortran} or APL\cite{APL}.
The result is a unique language that is applicable to wide array
	of applications.



To give you a taste of the language, in this program we approximate
\texttt{pi} using Monte Carlo integration

\begin{verbatim}
def f(a :: Real, b :: Real) :: Bool {
	return a*a + b*b < 1;
}
res :: Integer = zip(as, bs).count(f) / n;
\end{verbatim}

this would be translated into the map/reduce operations of

\begin{verbatim}
t1 = map(f, zip(as, bs));
count = 0;
reduce((x) => count += x, t1);
res = count / n;
\end{verbatim}

The compiler would lower the above into an IR representation,
maintaining the map/reduce operations. This is discussed 
in more detail later in the report.


The language is purely functional --- disallowing any side effects.
It is also statically typed --- with type anotation provided by the user.
Due to time constraints, it lacks many features such as conditionals, but 
	one can go around that limitation using simple bit operations.


